Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
proofread introduction

In this paper we aim to analyze polling data related to the 2024 election, with a particular emphasis on the swing states. By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions. Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states. Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election. We use Bayesian regression model to create a model to predict the percent of population that will support in the swing states based on pollster, the number of days since Kamala Harris joined the race. More on the model and the variables, in the Data and Model sections.

We noticed in most predictions of the presidential candidate they are usually based on one survey or pollster and they try to gather data from the one poll to predict the the future President. However, through looking at the results from different polls we are able to get a better, more accurate prediction. We also look at over state polls which allows us to focus on the swing states. Swing states are states that change sides every election. The 2 major parties that fight for power are the Republicans and the Democrats. Some states are always voting for one party regardless of the candidate so, the swing states are the once that tell us more about who might win the election. The gap we are trying to cover is predicting the future President of USA by looking at different polls, while being as effecient as possible by looking only at the swing states.

To go a little more in-depth into what we actually did, we have to first discuss our estimand. Our estimand, the parameter we aim to study, is the predicted support for Kamala Harris in the swing states (measured in %). We do this by creating Baysian model with 3 parameters; state, pollster, and says since Kamala Harris was announced as candidate. We have the priors set with a normal distribution centered at 0, with a standard deviation of 2.5. We expect that because the swing states are states that don't have a party that they dominantly vote for, the expected support will be around 50%. We also make other graphs in the Data section, which helps us understand the values themselves. 
From the result section we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%. Based on this we predict that Arizona, and Georgia will vote Donald trump and the rest of the states will vote Kamala Harris.

Predicting who will become president is important because it helps people understand and prepare for future changes in the country’s policies and priorities. A new president can shape laws, affect the economy, influence international relationships, and address key issues like healthcare, education, and climate change. Accurate predictions allow citizens, businesses, and governments to plan and adapt. For example, if polls indicate a particular candidate might win, groups can start thinking about how their plans might align with the candidate's proposed policies. It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy.

The paper is structured with the introduction in the beginning followed by the Data section wich takes a look at the methodology of the paper and how the data was gathered. We explain the variables we isolated in order for this analysis to be complete. Then it is followed by the Model section where we introduce the model and explain the graphs of related models. The result sections will cover the model summary and summary table of the predictions for each of the swing states. The last section is the discussion section where we talk about what the results means and we give our final conclusion and weaknesses. We also give a final word about who will win the upcoming election. We also have additional appendices, where we talk about the methodology of the polls and how they compare. The second appendix talks about a survey we designed with reasons and disadvantage. The third appendix just shows the models in more depth.

@@

# Model

@@

## Model Results

@@

# Discussion

>>>>

In this paper we aim to analyze polling data related to the 2024 election, with a particular emphasis on the swing states. By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions. Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states. Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the opinions shaping the 2024 presidential election. Using Bayesian regression modeling, we predict the percent of supprt for Kamala Harris in swing states based on pollster and the number of days since she announced her candidacy. Details on our data, variables, and model are discussed further in the Data and Model sections.

Traditional election forecasts often rely on single polls or sources, but our approach combines results from multiple polls to provide a more accurate prediction. By focusing on swing states, which historically alternate between parties, we capture the areas most likely to influence the election outcome. The 2 major parties that are in competition are the Republicans and the Democrats. Some states exhibit a consistent pattern of voting for a particular party, regardless of the candidate. In contrast, swing states are more variable, providing key insights into who might win the election. Our goal is to predict the outcome of the U.S. presidential election by synthesizing data from multiple polls, focusing efficiently on the indecisive swing states.

To delve further into our approach, we introduce our estimated - the parameter of interest- which is the predicted support for Kamala Harris in the swing states (measured in %). We estimate this by creating Baysian model with 3 parameters; state, pollster, and days since Harris’s candidacy announcement. We set priors using a normal distribution centered at 0, with a standard deviation of 2.5. Given that swing states tend to lack a dominant party preference, the expected support will be approximately 50%. We further explore these variables and their meanings other graphs in the (@sec-data). From our analysis in (@sec-results) we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%. Based on this we predict that Arizona, and Georgia will vote Donald trump and the rest of the states will vote Kamala Harris.

Accurate election predictions are important as they help individuals, businesses, and governments anticipate potential shifts in policy, the economy, and international relations. A new president can shape laws, affect the economy, influence international relationships, and address key issues like healthcare, education, and climate change. Knowing which candidate is likely to win enables groups to consider how their plans might align with anticipated policies, encouraging public engagement and participation in the democratic process. It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy in the future.

The paper is structured as follows: the Data section (@sec-data) covers our methodology and data collection process, detailing the variables used in the analysis. The Model section (@sec-model) introduces our Bayesian approach and includes related visualizations. The Results section (@sec-results) provides a model summary and predicted support percentages for each swing state. In the Discussion section (@sec-discussion), we interpret our findings, present final predictions, and discuss limitations, weaknesses and next steps.  We also give a final word about who will win the upcoming election. We also have additional appendices, where we talk about the methodology of the polls and how they compare. The second appendix talks about a survey we designed with reasons and disadvantage. The third appendix shows the Posterior Predictive check of our model.

@@

# Model {#sec-model}

@@

## Model Results {#sec-results}

@@

# Discussion {#sec-discussion}
